---
title: 'Reproducible Research Coursera: Peer Assessment 1'
---

1. Code for reading in the dataset and/or processing the data

```{r}
unzip ("C:/Users/desrjus/Documents/R Files/Coursera/RepData_PeerAssessment1/activity.zip")

activityData <- read.csv("C:/Users/desrjus/Documents/R Files/Coursera/RepData_PeerAssessment1/activity.csv")
summary(activityData)
```

2. Histogram of the total number of steps taken each day

```{r}
steps.per.day <- aggregate(steps ~ date, activityData, sum, na.rm=T)
hist(steps.per.day$steps)
```

3. Mean and median number of steps taken each day

```{r}
mean.steps.per.day <- mean(steps.per.day$steps)
mean.steps.per.day
median.steps.per.day <- median(steps.per.day$steps)
median.steps.per.day
```

4. Time series plot of the average number of steps taken

```{r}
steps.per.interval <- aggregate(steps~interval, data=activityData, mean, na.rm=T)
plot(steps~interval, data=steps.per.interval, type="l")
```

5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
max.steps.interval <- steps.per.interval[which.max(steps.per.interval$steps),]$interval
max.steps.interval
```

6. Code to describe and show a strategy for imputing missing data

```{r}
total.values.missing <- sum(is.na(activityData))
total.values.missing
```
- Number of total missing values
```{r}
get.mean.steps.per.interval <- function(interval){
  steps.per.interval[steps.per.interval$interval==interval,]$steps
}

activityDataFull <- activityData 
for(i in 1:nrow(activityDataFull)) {
    if(is.na(activityDataFull[i,]$steps)) {
      activityDataFull[i,]$steps <- get.mean.steps.per.interval(activityDataFull[i,]$interval)
    }
}
activityDataFull$steps <- as.numeric(activityDataFull$steps)
```
- New dataset with missing data filled in

7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}
library(tidyverse)
total.steps.per.day.full <- aggregate(steps ~ date, data=activityDataFull, sum)
hist(total.steps.per.day.full$steps)
```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
library(lubridate)
activityDataFull$date <- ymd(activityDataFull$date)
activityDataFull$day <- weekdays(activityDataFull$date)
for(i in 1:nrow(activityDataFull)) {
  if (activityDataFull[i,]$day %in% c("Saturday","Sunday")) {
    activityDataFull[i,]$day <- "weekend"
  }
  else{
    activityDataFull[i,]$day <- "weekday"
  }
}
steps.by.day <- aggregate(activityDataFull$steps ~ activityDataFull$interval + activityDataFull$day, activityDataFull, mean)

names(steps.by.day) <- c("interval", "day", "steps")
library(lattice)
xyplot(steps~interval | day, steps.by.day, type="l", layout=c(1,2),
       xlab="Interval", ylab="Number of Steps")
```

9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

- see above

